Mining of Keystroke and Mouse Dynamics to Increase the Engagement of Students with Programming Assignments

### 2018-04-21 paper.tex

    \end{tabular}
  \end{table}
  %
  \begin{table*}[h!tb]
\centering
\caption{Parameters used in the classification algorithms.}
\label{tab_clas}
\begin{tabular}{ | l | p{10cm} | }
  \hline
  Algorithm & Parameters \\
  \hline
  J48 & Confidence threshold for pruning = 0.25; minimum instances per
        leaf = 2; folds for reduced error pruning = 3 \\
  Neural network & Learning algorithm: Back-propagation; hidden layers
                   = 2, hidden layer size = 20 neurons; learning rate
                   = 0.6; momentum = 0.6; training iterations = 200 \\
  \hline
\end{tabular}
\end{table*}
%
As part of the data mining process a canonical genetic algorithm was used for feature
selection with the parameters used shown in Table \ref{tab_ga}. These
parameters were set heuristically, and were the same for selecting the
features of all classification algorithms. % Mario check this - JJ
In the end, the subset for the frustration classifier includes only 11 features, and 
the subset for the boredom classifier consists of only 13 features. 
%
% It's not clear what's the fitness function here or what it really does. Maybe we could expand this part of the work, test different evolutionary algorithms and write a paper on feature selection using evolutionary algorithms for EvoStar - JJ
%
The selected classifiers where implemented using Rapid Miner, with the
  parameters shown in Table \ref{tab_clas}.

@@

Table \ref{tab_performance} shows the performance of each of the four
classifiers together
with the $\kappa$ coefficients. A 10-Fold cross validation was used.
The accuracies and $\kappa$ coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}.
In a fixed-text method, subjects are asked to write specific texts.
The results obtained with this method are satisfactory, considering that
learners were writing a program, a task comparable with free-text writing,
where there is no restriction on what has to be written.
In the case of the $\kappa$ coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of $\kappa$ were close to 0.5, is important to consider the values of $\kappa$
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted $\kappa$
statistic for the kind of problem. The artificial neural network (ANN)
gives the highest
accuracy in the classification of frustration.
%
\begin{table}[h!tb]
\centering
\caption{Parameters used in the evolutionary algorithm for feature
  weight optimization.}
\label{tab_ga_w}
    \begin{tabular}{ | l | l | }
      \hline
      Parameter & Value \\
        \hline
      Population size & 10 \\
      Generations & 30 \\
      Selection & Tournament \\
      Tournament size & population size / 4 \\
      Crossover & Uniform \\
      Crossover probability & 0.5 \\
      Mutation rate & $1 \over 39$ \\
      Minimum number of features & 5 \\
        \hline
    \end{tabular}
  \end{table}
% 
\begin{table*}[!t]
\centering
\caption{Parameters used for the different classification algorithms
  used for recognizing the Flow/engaged state.}
\label{tab_params_ml}
    \begin{tabular}{ | l | p{10cm} |  }
\hline
      Algorithm & Parameters \\
      \hline 
      Random Forest &  number of trees = 10, 
criterion = gain ratio, maximal depth = 20, 
                      confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2 \\
      J48 &  confidence = 0.79, minimal leaf size=2 \\
      kNN & $k$ = 4 \\
      Gradient boosted trees & number of trees = 20,  maximal depth = 5, minimum rows = 10, 
minimum split 
improvement = 0.0, number of bins = 20, 
                               learning rate = 0.1, sample rate = 1.0 \\
      Bayes Kernel & laplace correction = True, estimation mode = greedy, 
                     minimum bandwidth = 0.1, number of kernels = 10 \\
      Deep learning & activation = Rectifier, hidden layer sizes =
                      [50, 50], epochs = 10 \\
      \hline
    \end{tabular}
  \end{table*}

Table \ref{tab_perf_flow} shows the performance for the additional classifiers
evaluated specifically for the Flow/engaged affective state. The goal was to
increase the performance achieved earlier. Again a ten-
fold cross-validation was conducted. In this case, instead of using an
evolutionary algorithm to optimize feature selection, it was used for 
optimizing a weight vector. 
      The feature-weight vector assigned a weight to each one of the
      features, and the parameters that have been used in the
      evolutionary algorithm are shown in Table \ref{tab_ga_w}.
      % The
% parameters used for the evolutionary optimization algorithm are the following:
% Population size = 10, generations = 30, selection = tournament, tournament size = 0.25 (fraction of the population),  initialize probability = 0.5,  mutation
%       probability = 1/number of attributes,  crossover probability =
%       0.5.
      % Please Mario check these parameters and how they have put to
      % the table; I don't know what's the "initialization
      % probability" - JJ
%
      The parameters used in this case for the classification
      algorithms, and the algorithms used are shown in Table \ref{tab_params_ml}
% used for the Random Forest algorithm: number of trees = 10, 
% criterion = gain ratio, maximal depth = 20, 
% confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2.  The
% parameters used for are the J48 are the following:  confidence = 0.79, minimal
% leaf size=2. For the kNN algorithm, the value k = 4 was used. 
% For gradient boosted
% trees: number of trees = 20,  maximal depth = 5, minimum rows = 10, 
% minimum split 
% improvement = 0.0, number of bins = 20, 
% learning rate = 0.1, sample rate = 1.0.  Bayes
% Kernel: laplace correction = True, estimation mode = greedy, 
% minimum bandwidth = 0.1, number of kernels = 10.  Finally for deep learning: activation = Rectifier, hidden layer sizes = [50, 50], epochs = 10.

>>>>

    \end{tabular}
  \end{table}
  %
  \begin{table*}[h!tb]
\centering
\caption{Parameters used in the classification algorithms.}
\label{tab_clas}
\begin{tabular}{ | l | p{10cm} | }
  \hline
  Algorithm & Parameters \\
  \hline
  J48 & Confidence threshold for pruning = 0.25; minimum instances per
        leaf = 2; folds for reduced error pruning = 3 \\
  Neural network & Learning algorithm: Back-propagation; hidden layers
                   = 2, hidden layer size = 20 neurons; learning rate
                   = 0.6; momentum = 0.6; training iterations = 200 \\
  \hline
\end{tabular}
\end{table*}
%
As part of the data mining process a canonical genetic algorithm was used for feature
selection with the parameters used shown in Table \ref{tab_ga}. These
parameters were set heuristically and were the same for selecting the
features of all classification algorithms. The population size was changed
mainly because some of the algorithms are more computationally demanding. The
genetic algorithm is included as a RapidMiner operator. In the algorithm, each
individual consists of a binary list indicating if a feature will be used or
not. The mutation operator simply changes the switch with the probability
indicated in the mutation-probability parameter.  When the population is
initialized each switch has 0.5 probability of being "on".

% Mario check this - JJ
% OK, added another sentence - Mario 
% Added a mini explanation of the GA. - Mario

In the end, the subset for the frustration classifier includes only 11 features, and 
the subset for the boredom classifier consists of only 13 features. 
%
% It's not clear what's the fitness function here or what it really does. Maybe we could expand this part of the work, test different evolutionary algorithms and write a paper on feature selection using evolutionary algorithms for EvoStar - JJ
%
The selected classifiers where implemented using Rapid Miner, with the
  parameters shown in Table \ref{tab_clas}.

@@

Table \ref{tab_performance} shows the performance of each of the four
classifiers together
with the $\kappa$ coefficients. A 10-Fold cross validation was used.
The accuracies and $\kappa$ coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}.
In a fixed-text method, subjects are asked to write specific texts.
The results obtained with this method are satisfactory, considering that
learners were writing a program, a task comparable with free-text writing,
where there is no restriction on what has to be written.
In the case of the $\kappa$ coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of $\kappa$ were close to 0.5, is important to consider the values of $\kappa$
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted $\kappa$
statistic for the kind of problem. The artificial neural network (ANN)
gives the highest
accuracy in the classification of frustration.
%
\begin{table}[h!tb]
\centering
\caption{Parameters used in the evolutionary algorithm for feature
  weight optimization.}
\label{tab_ga_w}
    \begin{tabular}{ | l | l | }
      \hline
      Parameter & Value \\
      \hline
      Initialize probability & 0.5 \\
      Population size & 10 \\
      Generations & 30 \\
      Selection & Tournament \\
      Tournament size & population size / 4 \\
      Crossover & Uniform \\
      Crossover probability & 0.5 \\
      Mutation rate & $1 \over 39$ \\
      Minimum number of features & 5 \\
        \hline
    \end{tabular}
  \end{table}
% 
\begin{table*}[!t]
\centering
\caption{Parameters used for the different classification algorithms
  used for recognizing the Flow/engaged state.}
\label{tab_params_ml}
    \begin{tabular}{ | l | p{10cm} |  }
\hline
      Algorithm & Parameters \\
      \hline 
      Random Forest &  number of trees = 10, 
criterion = gain ratio, maximal depth = 20, 
                      confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2 \\
      J48 &  confidence = 0.79, minimal leaf size=2 \\
      kNN & $k$ = 4 \\
      Gradient boosted trees & number of trees = 20,  maximal depth = 5, minimum rows = 10, 
minimum split 
improvement = 0.0, number of bins = 20, 
                               learning rate = 0.1, sample rate = 1.0 \\
      Bayes Kernel & laplace correction = True, estimation mode = greedy, 
                     minimum bandwidth = 0.1, number of kernels = 10 \\
      Deep learning & activation = Rectifier, hidden layer sizes =
                      [50, 50], epochs = 10 \\
      \hline
    \end{tabular}
  \end{table*}

Table \ref{tab_perf_flow} shows the performance for the additional classifiers
evaluated specifically for the Flow/engaged affective state. The goal was to
increase the performance achieved earlier. Again a ten-
fold cross-validation was conducted. Again, an evolutionary algorithm was used for feature with the parameters shown in Table \ref{tab_ga_w}.
      % The
% parameters used for the evolutionary optimization algorithm are the following:
% Population size = 10, generations = 30, selection = tournament, tournament size = 0.25 (fraction of the population),  initialize probability = 0.5,  mutation
%       probability = 1/number of attributes,  crossover probability =
%       0.5.
      % Please Mario check these parameters and how they have put to
      % the table; I don't know what's the "initialization
      % probability" - JJ
     % Is the probability a feature has to be "on" (used by the algorithm)
     % initially.
% Does that mean that you make probablity = 0.5? Still don't
% understand, but I have added to the table - JJ
      The parameters used in this case for the classification
      algorithms, and the algorithms used are shown in Table \ref{tab_params_ml}
% used for the Random Forest algorithm: number of trees = 10, 
% criterion = gain ratio, maximal depth = 20, 
% confidence = 0.25, minimal gain = 0.1, minmail leaf size = 2.  The
% parameters used for are the J48 are the following:  confidence = 0.79, minimal
% leaf size=2. For the kNN algorithm, the value k = 4 was used. 
% For gradient boosted
% trees: number of trees = 20,  maximal depth = 5, minimum rows = 10, 
% minimum split 
% improvement = 0.0, number of bins = 20, 
% learning rate = 0.1, sample rate = 1.0.  Bayes
% Kernel: laplace correction = True, estimation mode = greedy, 
% minimum bandwidth = 0.1, number of kernels = 10.  Finally for deep learning: activation = Rectifier, hidden layer sizes = [50, 50], epochs = 10.
